---
title: "DM Final Project_Logistic regression"
author: "Jianghui Wen"
date: "February 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Data
```{r}
dataPath <- "/Users/cherish/Desktop/Data_Mining"
uns.training <- read.csv(file=paste(dataPath, "uns_training.csv", sep="/"), header=TRUE)
uns.testing <- read.csv(file=paste(dataPath, "uns_testing.csv", sep="/"), header=TRUE)
```

```{r}
library(caret)
library(nnet)
library(MASS)
library(glmnet)
library(grid)
library(gridExtra)
library(plyr)
```

Fix the class name, make it consistant in both train and test dataset
```{r}
uns.training$UNS <- revalue(uns.training$UNS, c("very_low"="Very Low", "Low"="Low","Middle"="Middle","High"="High"))
head(uns.training)

```

Check the colinearity
```{r}
pairs(uns.training[,1:5])
```

#2. Multinomial Logistic Regression
```{r}
uns.multinomial <- multinom(UNS ~ STG+SCG+STR+LPR+PEG, data=uns.training, Hess= TRUE, method="class")
summary(uns.multinomial)
```

Using step() to select variables to build a good model
or calculate p values to select significant varaibles using the following 
ctable <- coef(summary(uns.polr))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p value" = p)
```{r}
step(uns.multinomial)
```

The result shows a good fitted model including predictors SCG,STR,LPR, PEG, with AIC=77.5996

```{r}
uns.multinomial.model <- multinom(UNS~SCG+STR+LPR+PEG, data=uns.training, Hess= TRUE, method="class")
summary(uns.multinomial.model)
```

Hold out for multinomial logistic regression 
```{r}
uns.multinomial.test <- predict(uns.multinomial.model, newdata= uns.testing, type="class")
round(prop.table(table(uns.testing$UNS, uns.multinomial.test)),2)
cm.table <- data.frame(c(26,0,0,0),c(0,45,1,0),c(0,1,31,0),c(0,0,2,39))
rownames(cm.table) <- c("Very Low","Low","Middle","High")
colnames(cm.table) <- c("Very Low","Low","Middle","High")
grid.table(cm.table)
```

#3.Proportional Odds logistic regression
```{r}
UNS.ordered <- ordered(uns.training$UNS, levels=c("Very Low","Low","Middle","High"))
uns.polr <- polr(UNS.ordered~STG + SCG + STR + LPR + PEG, data=uns.training, Hess= TRUE, method="logistic")
summary(uns.polr)
```

Using step() to select variables
```{r}
step(uns.polr)
```

The good fitted model including predictors SCG,STR,LPR,PEG, with AIC= 75.09955
```{r}
uns.polr.model <-  polr(UNS.ordered ~ SCG + STR+ LPR + PEG, data = uns.training, Hess = TRUE, method="logistic")
summary(uns.polr.model)
```

Holdout for proportional odds logistic regression 
```{r}
uns.polr.test <- predict(uns.polr.model, newdata= uns.testing, type="class")
```

Confution matrix for polr.model
```{r}
table(uns.testing$UNS, uns.polr.test)
round(prop.table(table(uns.testing$UNS, uns.polr.test)),2)
cm.dataframe <- data.frame(c(26,0,0,0), c(0,45,1,0),c(0,1,32,0),c(0,0,1,39))
rownames(cm.dataframe) <- c("Very Low","Low","Middle","High")
colnames(cm.dataframe) <- c("Very Low","Low","Middle","High")
grid.table(cm.dataframe)
```

#4. Compare the two models using Log Loss function
```{r}
log.loss <-  function(actual, pred){
    eps <- 1e-15
    if (!is.matrix(pred)) pred <- t(as.matrix(pred))
    if (!is.matrix(actual)) actual <- t(as.matrix(actual))
    nr <- nrow(pred)
    pred <- apply(pred, c(1,2), function(x) max(min(x, 1-10^(-15)), 10^(-15)))
    score <- -sum(actual*log(sweep(pred, 1, rowSums(pred), FUN="/")))/nr
    return(score)
}

response.matrix <- function(Y)
{
    y_res <- matrix(NA,nrow = 0,ncol = 4)
    for (i in 1:length(Y))
    {
        if (Y[i] == "Very Low")
            y_res <- rbind(y_res,c(1,0,0,0))
        if (Y[i] == "Low")
            y_res <- rbind(y_res,c(0,1,0,0))
        if (Y[i] == "Middle")
            y_res <- rbind(y_res,c(0,0,1,0))
        if (Y[i] == "High")
            y_res <- rbind(y_res,c(0,0,0,1))
    }
    return (y_res)
}
```

```{r}
multinomial.prediction <- predict(uns.multinomial.model, newdata= uns.testing, type="probs")
log.loss(response.matrix(uns.testing$UNS), multinomial.prediction[,c(4,2,3,1)])

```
*Log Loss value for multinomial model is: 0.06869159*

```{r}
polr.prediction <- predict(uns.polr.model, newdata= uns.testing, type="probs")
log.loss(response.matrix(uns.testing$UNS), polr.prediction)
```
*Log loss value of proportional odds is: 0.05271212*

#Response time
```{r}
#multinomial model
system.time(uns.multinomial <- multinom(UNS ~ STG+SCG+STR+LPR+PEG, data=uns.training, Hess= TRUE, method="class"))
system.time(step(uns.multinomial))

#polr model
system.time(uns.polr <- polr(UNS.ordered~STG + SCG + STR + LPR + PEG, data=uns.training, Hess= TRUE, method="logistic"))
system.time(step(uns.polr))
```

response time:              user  system   elapsed 
uns.multinom                0.030   0.004    0.051
step(multinom)              0.263   0.037    0.332 
uns.polr                    0.035   0.003    0.047 
step(Polr)                  0.295   0.018    0.356 

####microbenchmark
```{r}
library(microbenchmark)
microbenchmark(uns.multinomial, step(uns.multinomial), uns.polr, step(uns.polr))
```

